### 2015-04-22 thesis/thesis/chap5/chap5.tex
changes

\chapter{On Probabilistic Group Social Commitments}\label{cha:PCTLKC+}

In this chapter\footnote{Part of the results presented in this chapter, namely PCTL$^{\textrm{kc+}}$ logic, has been published in SoMet\_14 \cite{Sultan2014c}. The results of model checking PCTL$^{\textrm{kc+}}$ have been sumbitted to the Engineering Applications of Artificial Intelligence journal \cite{Sultan2014d}}, we improve and extend the work presented in Chapter \ref{cha:PCTLKC} by refining the probabilistic logic of knowledge and commitment (PCTL$^{kc}$) and then extending the refined logic further by operators for group knowledge and group commitment. In this respect, we define a semantics for the group social commitment operator and integrate it into the resulting logic. The developed logic is called the new probabilistic logic of knowledge and commitment (PCTL$^{\textrm{kc+}}$). Finally, we introduce a new formal verification technique that considers the new group modalities and implement it on top of the PRISM model checker.

\section{Introduction} \label{Introduction}

One of the major challenges in building complex software products
such as Multi-Agent Systems (MASs) is to advance error detection
at early stages of their life-cycles. %The social context of interacting agents in MASs has gained significance recently.
MASs community has witnessed an important shift in defining the semantics of ACLs from the so-called mental approaches that is hard to verify \cite{Singh2000} to social approaches which exploit observable and verifiable social commitments. However, the increasing demand to use social commitments as a means of communication among interacting parties \cite{Baldoni2014,Chesani2013,Akin2013} requires reasoning about and verifying the relationship between social commitments and some other systems' aspects such as agents' knowledge and uncertainty especially in the case of having group-commitment scenarios. In addition to verifying the interactions between social commitments and knowledge in the presence of uncertainty, the ultimate objective of this chapter is to verify the interactions between the two elements when the scope of interacting agents goes beyond the common agent-to-agent (i.e., one-to-one) scheme.

In order to effectively capture and express the interactions between individual and group social commitments and knowledge in probabilistic MASs, we propose a new modal logic called the new probabilistic logic of knowledge and commitments PCTL$^{\textrm{kc+}}$ which is built by combining a consistent logic of knowledge and commitment CTLKC$^+$ \cite{Al-Saqqar2014a} with a well established probabilistic temporal logic PCTL \cite{Hansson1994}. The resulting logic is extended further to accommodate operators for the group knowledge and group commitments.

At present, there is a relatively large gap in addressing the
concepts of knowledge and social commitments simultaneously in
MASs, especially with the presence of uncertainty. Existing
approaches that address the interaction between knowledge and
social commitments either limit the scope of interacting
agents to the widely used one-to-one commitment scheme and ignore the uncertainty aspect of MASs \cite{Al-Saqqar2014a}, or adopt a different kind of commitments called ``internal commitment'' rather than the ``social commitments'' that we consider in this thesis \cite{Schmidt2004}.
Furthermore, although the notion of ``group'' is important in the multi-agent community \cite{Sanchez-Anguix2013,YuW2010}, group social
commitments has not been formalized and verified yet. As knowledge
and social commitments influence each other in many real world
applications \cite{Al-Saqqar2014a}, their interactions need to be
reasoned about and verified in a systematic manner. As we said before,
uncertainty in MASs may arise due to imperfect information about
the environment in which agents interact. Besides, in some situations, it happens that even if there is some state of affairs (i.e., content of a commitment) that an agent wants to bring about, its actions might not reliably drive the state of affairs into the desired state \cite{Sultan2014a}. Consequently, commitments themselves become stochastic and the degree to which the commitment can be satisfied is not always
guaranteed.

To motivate our study of representing and verifying the
interaction between individual and group knowledge and social
commitments in probabilistic MASs when taking into account
one-to-many commitments, let us consider the following simple
example. A professor teaching an engineering course with a
capacity of $20$ students. Various scenarios could happen within
this context.

\begin{itemize}
\item \verb"Scenario 1": while the professor was explaining some
new concepts in the course, one of the students asked the
professor to provide him with more material regarding these
concepts. The professor then promised to email the student some
references the next day. This promise can be considered as a
commitment from the professor towards the student to provide him
with extra material for the new concepts. \item \verb"Scenario 2":
in the lecture before the mid-term exam, students requested the
professor to exclude some material and shorten the duration of the
exam accordingly. At the end of the lecture, the professor agreed
to exclude some parts of the covered material and to make the exam
one hour long.  This agreement can be considered as a commitment
from the professor to the group of students who are registered in
this course. Right after the class, the professor posted in the
course web site an update confirming what they agreed on.
\end{itemize}
In the first scenario, obviously both the professor and student
are aware of the commitment (sending extra materiel). However, in
the second scenario, every student registered in the course, and
not necessarily present during the lecture,  has to know about
this agreement (commitment) to avoid wasting time studying
excluded parts. In fact, because of some unexpected factors like
absence of the professor on the day of the exam, email delivery
failure, power outage, etc, there is no guarantee that these
commitments are  going to be surely fulfilled. Therefore, it is
important to have a logic system with the ability to not only
express the interaction between knowledge and social commitments,
but also to handle the concepts of knowledge and commitments
within the scope of a group when uncertainty matters. Once such a
logic is defined, it can be invested as the underlying logic for a
verification technique to verify some desirable and useful
properties expressing combinations of knowledge and social
commitments under uncertainty.

The work presented in this chapter can be seen as extension and continuation of the work presented in Chapter \ref{cha:PCTLKC} where reasoning about and verifying interactions between individual knowledge and social commitments in probabilistic MASs were first introduced. In Chapter \ref{cha:PCTLKC}, we proposed a probabilistic logic called PCTL$^{kc}$ whose expressiveness power allowed us to formulate combinations of the two concepts in the presence of uncertainty. PCTL$^{kc}$ logic was built by fusing two logics, namely PCTLC \cite{Sultan2013} and PCTLK \cite{Wan2013} using the
independent join technique \cite{Franceschet2004}. However, as pointed out in \cite{Al-Saqqar2014a}, the social accessibility relations given in \cite{Bentahar2012,El-Menshawy2013a}, have an over-specification problem, and consequently the PCTL$^{kc}$ logic suffers from some paradoxes as it adopts the aforementioned accessibility relations.

To elaborate, there exist some situations that are not desirable in real settings but with the use of PCTL$^{kc}$, they are valid. One major problem in PCTL$^{kc}$ is that agents commit everything they know to others, which brings the lack of privacy into being. Formally, this is represented by the following postulate:

\begin{itemize}
\item $K_i \varphi \Rightarrow C_{i \rightarrow j}\varphi$,  where $i \neq j$.
\end{itemize}

The validity of this postulate is based on the fact that by
establishing communication channels through which commitments are
supposed to be exchanged, the epistemic relation needed to define
the semantics of knowledge is also established. This is not
reasonable in open environments where agents are selfish. Another
problem is that agents commit everything known by others. That is,
when an agent knows that another agent knows something, the first
agent commits to bring about what the other agent knows, formally:

\begin{itemize}
\item $ K_i K_j \varphi \Rightarrow C_{i \rightarrow j} \varphi$ where $ i \neq j$.
\end{itemize}

Such a postulate should be avoided in MASs because it is not
realistic for an agent to commit for something that is out of its
capabilities. Moreover, this postulate can result in serious
circumstances if agent $j$ is malicious, so it can express
incorrect knowledge about the other agent, obliging it to
establish unwanted commitment.

On the other hand, we have some reasonable situations that should
be always valid but with the use of PCTL$^{kc}$ they can be
unsatisfied. One example in this respect is when agents should be always aware about the fulfillment of their own commitments.

\begin{itemize}
\item $Fu(C _{i\rightarrow j} \varphi) \Rightarrow K_i Fu(C
_{i\rightarrow j} \varphi)$ where $i \neq j$
\end{itemize}

It is realistic for this postulate to be valid because any agent should be aware of its fulfillment actions in order to prevent fulfilling the same commitment again and again. However, this postulate is not valid in PCTL$^{kc}$ because of the same reasons mentioned earlier. Consequently, PCTL$^{kc}$ fails to efficiently handle some practical situations in which knowledge and social commitments need to interact.

The problem is that one of the underlying logics of PCTL$^{kc}$, which
is PCTLC \cite{Sultan2013}, is built using the social accessibility relations given in \cite{Bentahar2012,El-Menshawy2013a}, which in fact over specifies and over constrains the concept of illocutionary communication. In a recent work, Al-Saqqar et al. \cite{Al-Saqqar2014a} have figured out that although the social accessibilities presented in \cite{Bentahar2012,El-Menshawy2013a} function perfectly when the concern is to model social commitments independently, they have some limitations when combined with the epistemic accessibility relations in the same model. The authors in \cite{Al-Saqqar2014a} modified the social accessibility relations proposed in \cite{Bentahar2012,El-Menshawy2013a} in order to prevent the unintended emergence of the epistemic accessibility relations from the social accessibility relations when the two accessibilities combined in the same model. Technically speaking, they relaxed the conditions upon which the social accessibility relations are established in order to decouple the social accessibility relations from the epistemic accessibility relations. The new definition of social accessibilities does no longer depend on the unshared variables but rather depends merely on the shared variables between the interacting agents as discussed in Chapter \ref{cha:background}. The new condition upon which a communication channel is established is stated below:

 $ s \approx_{i \rightarrow j} s' $ iff $ Var_i \cap Var_j \neq \emptyset $ such that $ \forall x \in Var_i \cap Var_j $ we have $ l_i^x(s) =
l_i^x(s') = l_j^x(s')$, where $\approx_{i\rightarrow j} \subseteq S \times S$ is the social accessibility relation. It has been proven that with the new social accessibilities, the resulting logic is consistent \cite{Al-Saqqar2014a}.

The work presented in this chapter differs from the one proposed in Chapter \ref{cha:PCTLKC} in the following points:

\begin{enumerate}
\item While the logic presented in Chapter \ref{cha:PCTLKC} suffers from
some paradoxes, the current work builds upon a consistent logic of
knowledge and commitment CTLKC$^+$ \cite{Al-Saqqar2014a} which
ensures having a paradox-free logic.

\item The new logic allows us to reason about commitments among
multiple agents instead of limiting the scope to merely two
agents. The concept of group knowledge is also integrated to the framework allowing us to reason about the knowledge in the case of group of agents.

\item In the current work, we generalize the model checking
technique that has been proposed in Chapter \ref{cha:PCTLKC} to fit the new group commitment operators as well.
\end{enumerate}

The contributions of this chapter are threefold. First, we present a new probabilistic logic called (PCTL$^{\textrm{kc+}}$) with expressiveness abilities to capture and represent the interactions between individual and group knowledge and social commitments.
Second, we introduce a formal verification technique for the probabilistic logic PCTL$^{\textrm{kc+}}$. The proposed technique is based on reducing
the problem of model checking PCTL$^{\textrm{kc+}}$ to the problem
of model checking PCTL. This is achieved through 1) advocating a
set of transformation rules that transform the
PCTL$^{\textrm{kc+}}$ model into a Markov Decision Process (MDP),
and then converting the obtained MDP into a DTMC using the notion
of ``adversary'' \cite{Forejt2011}; 2) reducing
PCTL$^{\textrm{kc+}}$ formulae into PCTL formulae based on a set
of formal reduction rules. Third, we implement our reduction model checking
technique on top of PRISM and apply it on a concrete case study,
namely the online shopping system \cite{Gomaa2011}. We then check
some system's properties written as PCTL$^{\textrm{kc+}}$ formulae
using the PRISM model checker by checking their corresponding PCTL
formulae. Figure \ref{fig:approach-PCTLKC+} depicts a schematic
view of our proposed framework.

\begin{figure}[t]
                \begin{center}
                \includegraphics[width=14cm, height=10cm]{chap5/img/approach-cha5-new.eps}
                \end{center}
                \caption{A schematic view of the probabilistic group social commitment approach} \label{fig:approach-PCTLKC+}
                \label{Stack}
                \end{figure}

\section{The New Probabilistic Logic of knowledge and Commitment (PCTL$^{\textrm{kc+}}$)} \label{sec:logic}

To overcome the inconsistency problem of PCTL$^{kc}$, we develop a new logic called the new probabilistic logic of knowledge and commitment (PCTL$^{\textrm{kc+}}$). To build PCTL$^{\textrm{kc+}}$, there are two obvious resources available in the literature: 1) the traditional temporal logics that have been developed for knowledge and social commitments independently or together such as CTLC \cite{Bentahar2012}, CTLK \cite{Lomuscio2007}, and CLTKC$^+$ \cite{Al-Saqqar2014a}; and 2) the existing probabilistic logics available in the literature such as PCTL \cite{Hansson1994}, PCLTK \cite{Wan2013}, and PCTLC \cite{Sultan2014a}. Unfortunately, none of these resources is perfectly suitable for the task. The former resource neglects the uncertainty aspects in MASs, while the
latter doesn't capture the interaction between knowledge and
social commitments. Therefore, we propose a solution that draws
upon both resources. In particular, we combine an existing
consistent logic of knowledge and commitment CTLKC$^+$
\cite{Al-Saqqar2014a} with a well established probabilistic
temporal logic PCTL \cite{Hansson1994}. Then, we extend the
resulting combined logic by adding new operators for the group knowledge
and group commitment.

Before going further, let us first describe a new probabilistic model over which PCTL$^{\textrm{kc+}}$ formulae can be interpreted. This model is an extension of the formalism of interpreted systems
\cite{Fagin1995} with the concepts of epistemic accessibility and
social accessibility relations. Concretely, the model of PCTL$^{\textrm{kc+}}$ is generated from combining two extended versions of the interpreted systems formalism. These extended formalisms are the extended version introduced in \cite{Halpern2003,Wan2013} and a modified version of the extended version given in \cite{Bentahar2012,El-Menshawy2013a} due to Al-Saqqar et al. \cite{Al-Saqqar2014a}.

\begin{definition} [PCTL$^{\textrm{kc+}}$ Model] ~\label{dfn:Models}

\noindent Given a set of atomic propositions $\Phi_p = (p,q,r, \ldots)$ and a set of agents $\texttt{Agt}=\{1,\ldots,n\}$,  the model $\mathfrak{M_3}=(S,\textbf{P},I,\sim_1, \ldots
,\sim_n,\{\approx_{i \rightarrow j}\}_{{(i,j)}\in \texttt{Agt}^2},\nu)$ is a tuple where:
%
\begin{itemize}
\item  $S \subseteq L_1 \times \ldots \times L_n$ is a countable set of all reachable global states of the system. A state $s$ is reachable iff there exists a sequence of transitions from an initial state to $s$ in which the probability of each transition is greater than $0$.

\item  $I \in S$ is an initial global state for the system.

\item  $\textbf{P}:S\times S\rightarrow [0,1]$ is a total transition probability function defined as $\textbf{P}(s, s')=\tau(s,a^{s \rightarrow s'}, s')$ iff there exists a joint action $a=(a_1,\ldots,a_n) \in ACT$ such that\\
     $\sum_{i \in \texttt{Agt}} \tau_i(l_i(s),a^{l_i(s)\rightarrow l_i(s')},l_i(s')) > 0$ and $\sum_{s' \in S} \textbf{P}(s,s') =1$ for all $s \in S$.

\item  $\sim_i \subseteq S \times S$ is the epistemic accessibility relation for the agent $i$, such that for two global states $s$ and $s'$, we have: $s \sim_i s'$ iff $l_i(s)=l_i(s')$.

\item For each pair $(i,j) \in \texttt{Agt}^2$, $\approx_{i\rightarrow j} \subseteq S \times S$ is the social accessibility relation which is defined as follows: $ s \approx_{i\rightarrow j} s' $ iff $ Var_i \cap Var_j \neq \emptyset $ such that $ \forall x \in Var_i \cap Var_j $ we have $ l_i^x(s) = l_i^x(s') = l_j^x(s')$.

\item  $\nu : S \rightarrow 2 ^{\Phi_p} $ is a valuation function.

\end{itemize}
\end{definition}
%

The new model $\mathfrak{M_3}$ differs from the model $\mathfrak{M_2}$,  presented in Chapter \ref{cha:PCTLKC}, in one particular point which is the social accessability relation. While $\mathfrak{M_2}$ uses the social accessibility relation $\sim_{i\rightarrow j}$ that has been introduced in \cite{Bentahar2012,El-Menshawy2013a}, the new model adopts the one $\approx_{i\rightarrow j}$ proposed in \cite{Al-Saqqar2014a} in order to overcome the over-specification problem appeared in $\sim_{i\rightarrow j}$.

\subsection{Syntax of PCTL$^{\textrm{kc+}}$} \label{def:syntax-PCTLKC+}

\begin{definition}[PCTL$^{\textrm{kc+}}$ syntax]\label{def:syntax-PCTLKC+}
Let $\Phi_p=\{p,q, \dots\}$ be a set of atomic propositions, and $\texttt{Agt}=\{1, \dots, n\}$ be a set of agents. The syntax of PCTL$^{\textrm{kc+}}$, which is a combination of PCTLK \cite{Wan2013} and PCTLC \cite{Sultan2014a,Sultan2013} augmented with further operators for the group knowledge, is given by the following grammar:

\vspace{-0.5cm}
\begin{align*}
    \varphi & ::= p~|~\neg \varphi~|~\varphi \vee \varphi~|~\mathcal{K}~|~\mathcal{C}~|~ \mathbb{P}_{\bowtie k} (\psi)~|~\mathbb{P}_{\bowtie k}(\mathcal{K})|~\mathbb{P}_{\bowtie k}(\mathcal{C})\\
    \psi & ::=\bigcirc \varphi ~ | ~ \varphi ~U~ \varphi~|~ \varphi~ U^{\leq m} ~ \varphi \\
    \mathcal{K} & ::= K_i ~\varphi ~|~ E_{G} ~\varphi\\
    \mathcal{C} & ::= C_{i\rightarrow j} ~\varphi ~| ~ C_{i\rightarrow G} ~\varphi ~| ~Fu(C_{i\rightarrow j} ~\varphi) ~| ~Fu(C_{i\rightarrow G} ~\varphi)
\end{align*}
%
where;\\
\noindent -- $p\in\Phi_p$ is an atomic proposition\\
\noindent -- $i,j \in \texttt{Agt}$.\\
\noindent -- $G\subseteq \texttt{Agt}$.\\
\noindent -- $\mathbb{P}_{\bowtie k}$ is a probabilistic operator and $\bowtie \in\{<,\leq,>,\ge\}$.\\
\noindent -- $k\in [0,1]$ is a probability bound or threshold.\\
$m \in\mathbb{N}^+ $ is a positive integer number reflecting the maximum number of transitions needed to reach a certain state.\\
\noindent -- The Boolean connectives $\neg$ and $\vee$ are defined in the usual way.\\
\noindent -- $\varphi$ and $\psi$ are state and path formulae interpreted over the states and paths of $\mathfrak{M_3}$ respectively.\\
\noindent -- The modal connectives $\mathcal{K}$ and $\mathcal{C}$ stand for ``epistemic" and ``social" operators, respectively.\\

\end{definition}
\vspace{-0.5cm}

\noindent In this logic, formulae $\mathcal{K}$ are state formulae and used to express the epistemic properties through the operators; $K_i$ which stands for agent $i$ knows, $E_G$ which stands for everyone knows. Modal connectives $C_{i\rightarrow j}$ and $C_{i\rightarrow G}$ are called social formulae and stand for ``commitment'' from a debtor towards a single creditor, and ``commitment'' from a debtor to a group of creditors, respectively. Likewise, modal connectives $Fu(C_{i\rightarrow j})$ and $Fu(C_{i\rightarrow G})$ stand for ``fulfillment'' of the commitment $C_{i\rightarrow j}$ and ``fulfillment'' of the commitment $C_{i\rightarrow G}$, respectively. $\bigcirc, U$ and $U^{\leq m}$ stand for ``next time'', ``until'' and ``bounded until'' path modal connectives respectively.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Social Commitments Classification} \label{sec:commitment-classification}

Social commitments for agent communication have been always looked
at within the scope of one-to-one. However,
back to our motivating example, we realize that in addition to the
usual agent-to-agent scheme, there are certain situations where
group-agent commitments are needed. In this chapter, we are
interested to move beyond the scope of one-to-one and
investigate the case of committing to multiple agents. The idea of
investigating other schemes of social commitments rather than the
one-to-one commitment scheme seems to be both technically
interesting and intuitively appealing. In what follows, we
distinguish between two different flavors of social commitments,
namely basic (or individual) social commitment and group social
commitment.

\begin{definition} [Basic Social Commitment]~

A basic social commitment is an agreement between two agents
namely, \texttt{debtor} and \texttt{creditor} such that the
\texttt{debtor} engages towards the \texttt{creditor} to bring
about a certain property.
\end{definition}

This is the simplest form of social commitments and has long been
investigated in the literature. The commitment in this case can be
represented using the following operator: $C_{i\to j} ~\varphi$
where $i$ denotes the debtor, $j$ denotes the creditor, and
$\varphi$ denotes the content of the commitment. The fulfillment
of such a commitment is written as follows: $Fu(C_{i\to j}
~\varphi)$. However, as the common form of social commitments is
the basic social commitment, we can simply use ``social
commitments" to refer to ``basic (or individual) social
commitments".

\begin{definition} [Group Social Commitment]~

A group social commitment is an agreement between a
\texttt{debtor} and a group of \texttt{creditors} to bring about a
certain property.
\end{definition}

This kind of commitments indicates the involvement of multiple
agents in the same commitment. The creditor is a group of
independent agents that join together as a single party due to
their shared interests in the commitment at hand. A group social
commitment is represented using the following notation: $C_{i\to
G} ~\varphi$, where $i$ denotes the debtor, $G$ denotes a group of
creditors, and $\varphi$ denotes the content of the commitment.
The fulfillment of such a commitment is given by the notation
$Fu(C_{i\to G} ~\varphi)$. Technically, a group social commitment
can be seen as the conjunction of individual basic social
commitments from the debtor $i$ to each agent in the group of
creditors $G$. Formally, $C_{i\to G} ~\varphi \equiv
\bigwedge\limits_{j\in G} C_{i\rightarrow j} ~\varphi$. An
intuitive explanation of the operator $C_{i\to G} ~\varphi$ is as
follows: for a group social commitment to be held at a certain
state, the content of the commitment must be true at every
accessible state from the commitment state with respect to the
group. This implies that none of the group members could be
excluded from having all accessible states satisfy the content of
the commitment. Consequently, it is obvious that for a state to be
socially accessible from the commitment state with respect to the
group, it has to be socially accessible with respect to at least
one of the agents of the group. Therefore, we resolve the
accessibility problem resulted from having group commitments by
taking the union of the social accessibility relations of each
single agent in the group. This in turn leads us to define the
group social accessibility relation based on the social
accessibility relations presented in Definition \ref{dfn:Models}.

Let $G\subseteq \texttt{Agt}$ be a group of agents. We define the group social accessibility relation from the social accessibility relation $\approx_{i \to j}$ as follows:
\begin{definition} [Social Accessibility Relations for Group Social Commitment] \label{dfn: group social
relations}~
%
\begin{itemize}

\item $\approx_{i \rightarrow G}$ is the union of the social accessibility relations between agent $i$ and each agent in the group $G$: $\approx_{i \to G}=\bigcup\limits_{j \in G}\approx_{i\to j}$.

\end{itemize}

\end{definition}

\begin{figure}[t]
                \begin{center}
                \includegraphics[width=7cm, height=5cm]{chap5/img/group-sc.eps}
                \end{center}
                \caption{Accessibility relations for group social commitment} \label{fig:group-sc}
                \end{figure}

Notice that group social commitments have all the properties of basic social commitments with an additional constraint that they can involve more than two agents. However, a group social commitment involving only two agents is equivalent to the basic social commitment. Figure \ref{fig:group-sc} depicts the idea of group social accessibility ($G=\{j_1, j_2\}$).

\subsection{Group Knowledge}  \label{sec:Group Knowledge}
In this work, we limit the scope of group knowledge to the concept of ``Everyone Knows" introduced in \cite{Fagin1995}. ``Everyone Knows" is denoted by $E_G ~\varphi$ and means that everyone in the group $G$ knows $\varphi$. Technically, ``Everyone knows" can be seen as the conjunction of the individual knowledge of each agent in the group. Formally, $E_G ~\varphi \equiv \bigwedge\limits_{i\in G} K_i ~\varphi$.

Before we proceed to present the semantics of
PCTL$^{\textrm{kc+}}$, we need to define the epistemic
accessibility relation for $E_G ~\varphi$. Let $G\subseteq
\texttt{Agt}$ be a group of agents. We define the epistemic
accessibility relation for $E_G ~\varphi$ from the epistemic
accessibility relation $\sim_i$ as follows \cite{Wan2013}:\\

\begin{definition} [Epistemic Accessibility Relation for Everyone Knows] \label{dfn: epistemic relations}~
\begin{itemize}

\item $\sim_G^{E}$ is the union of group $G's$ accessibility
relations: $\sim_G^{E}=\bigcup\limits_{i \in G}\sim_i$.

\end{itemize}
\end{definition}

\subsection{Semantics of PCTL$^{\textrm{kc+}}$} \label{def:semantics-PCTLKC+}

Given a model $\mathfrak{M_3}=(S,\textbf{P},I,\sim_1, \ldots
,\sim_n,\{\approx_{i \to j}\}_{{(i,j)}\in \texttt{Agt}^2},\nu)$, then $(\mathfrak{M_3},s) \models \varphi$ states that ``a state $s$ in the model $\mathfrak{M_3}$ satisfies a state formula $\varphi$, $(\mathfrak{M_3},\pi) \models \psi$ means that ``a path $\pi$ in the model $\mathfrak{M_3}$ satisfies a path formula $\psi$, and
$(\mathfrak{M_3},s) \models \mathbb{P}_{\bowtie k}(\psi)$ means that
``a state $s$ in $\mathfrak{M_3}$ satisfies
$\mathbb{P}_{\bowtie k}(\psi)$ if the probability of taking a path
from $s$ that satisfies $\psi$ is in the interval specified by
$\bowtie k$''. When the model $\mathfrak{M_3}$ is clear from the
context, we simply write the satisfaction relation $\models$ as
follows: $s \models \varphi$ and $\pi \models \psi$. Furthermore,
we denote the number of socially accessible states $s'$ from a given state $s$ such that $s\approx_{i\to j}s'$ by $|s\approx_{i\to j}s'|$, and $s\approx_{i\to G}s'$ by $|s\approx_{i\to G}s'|$.
We also denote the number of epistemically accessible states $s'$ from a given state $s$ such that $s\sim_i s'$ by $|s\sim_i s'|$. Similarly, we
denote the number of states $s'$ that are accessible from a given
state $s$ through $\sim_G^E$ by $|s\sim_G^E s'|$.
Finally, we define $|s\models \varphi|$ as follows:

\begin{center}
$|s\models \varphi|=
\begin{cases}
1,~~~\textrm{if}~ s\models\varphi\\
0,~~~\textrm{otherwise}.
\end{cases}$
\end{center}

%
\begin{definition}[\textbf{Satisfaction}]\label{def:semantics-pctlkc+} Satisfaction of a PCTL$^{\textrm{kc+}}$ formula in the model $\mathfrak{M_3}$ is conductively defined as follows:

\noindent $s\models p~~~~~~~~~~~~~~~~~~~\emph{iff}~~p\in \nu(s);\\
s\models \varphi_1 \vee \varphi_2 ~~~~~~~~\emph{iff}~~s\models \varphi_1~\textrm{or}~s \models \varphi_2; \\
s\models \neg \varphi~~~~~~~~~~~~~~~~\emph{iff}~~s \nvDash \varphi;\\
s\models K_i \varphi  ~~~~~~~~~~~~~~ \emph{iff}~~ \forall s'\in S ~ \textrm{s.t.} ~~ s \sim_i s',  \ \textrm{we~have}~ s'\models \varphi; \\
s\models E_G ~\varphi ~~~~~~~~~~~~ \emph{iff}~~\forall s'\in S~ \textrm{s.t.} ~~ s\sim_G^E s', \ \textrm{we~have}~ s'\models \varphi; \\
s\models C_{i\rightarrow j} \varphi  ~~~~~~~~~~ \emph{iff} ~~\forall s' \in S~\textrm{s.t.}~~s \approx_{i \rightarrow j}s', \textrm{we~have}~s'\models K_i \varphi \wedge K_j \varphi;\\
s\models C_{i\rightarrow G} \varphi  ~~~~~~~~~ \emph{iff} ~~\forall s' \in S~\textrm{s.t.}~~s \approx_{i \rightarrow G}s', \textrm{we~have}~s'\models K_i \varphi \wedge E_G ~\varphi $;\\
$s\models Fu(C_{i\rightarrow j} \varphi)$ ~\emph{iff} there exists $ s' \in S $ such that $ s' \approx_{i \rightarrow j} s $ and $~s'\models C_{i\rightarrow j} ~\varphi$ or \\
$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$there exists $ s'' \in S $ and $ s'' \sim_i s $ such that $~s''\models Fu(C_{i\rightarrow j} ~\varphi)$ or \\
$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$there exists $ s'' \in S $  and $ s'' \sim_j s $ such that $~s''\models Fu(C_{i\rightarrow j} ~\varphi)$;\\
$s\models Fu(C_{i\rightarrow G} ~\varphi)$ ~\emph{iff} there exists $ s' \in S $ such that $ s' \approx_{i \rightarrow G} s $ and $~s'\models C_{i\rightarrow G} ~\varphi$ or \\
$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$there exists $ s'' \in S $ and $ s'' \sim_i s $ such that $~s''\models Fu(C_{i\rightarrow G} ~\varphi)$ or \\
$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~$there exists $ s'' \in S $  and $ s'' \sim_G^E s $ such that $~s''\models Fu(C_{i\rightarrow G} ~\varphi)$;\\
\noindent $\pi \models \bigcirc \varphi~~~~~~~~~~~~~~\emph{iff}~~\pi(1) \models \varphi; \\
\pi \models \varphi_1~U^{\leq m}~\varphi_2~~\emph{iff}~~\exists k \leq m~~\textrm{s.t.}~~ \pi(k) \models \varphi_2 ~\textrm{and}~\forall i < k, \pi(i) \models \varphi_1;\\
\pi \models \varphi_1 ~U~\varphi_2~~~~~~~\emph{iff}~~\exists m \geq 0~~\textrm{s.t.}~~\pi \models \varphi_1~U^{\leq m}~\varphi_2;$\\
\noindent $s \models \mathbb{P}_{\bowtie k} (\psi)~~~~~~~~~\emph{iff}~~Prob_s(\psi)\bowtie k$ where: $Prob_s(\psi)=Prob_s\{\pi \in \Pi(s)~|~\pi\models
\psi\};$\\

\noindent $\bullet$ For a probabilistic operator working on an epistemic formula, where the set of all accessible states from $s$ is our sample space and the set of events $\mathrm{F}$ is the set of states accessible from $s$ and satisfy the formula:

\begin{tabbing}
\noindent $s\models \mathbb{P}_{\bowtie k}(K_i~\varphi) $
    \ \ \ \ \ \= \emph{iff} \ $Prob(s \models K_i\varphi)$ $\bowtie k$ where: $Prob(s\models K_i\varphi)=\frac{\sum_{s \sim_i s'}|s'\models \varphi| }{|s \sim_i s'|};$
\end{tabbing}

\begin{tabbing}
 \noindent $s\models \mathbb{P}_{\bowtie k}(E_G~\varphi) $
    \ \ \ \ \ \= \emph{iff} \ $Prob(s \models E_G~\varphi)$ $\bowtie k$ where: $Prob(s\models E_G~\varphi)=\frac{\sum_{s \sim_G^E s'}|s'\models \varphi| }{|s \sim_G^E s'|};$

\end{tabbing}

\noindent $\bullet$  For a probabilistic operator working over a commitment formula,where the set of all accessible states from $s$ is our sample space and the set of events $\mathrm{F}$ is the set of states accessible from $s$
and satisfy the formula: \\

\noindent $s\models \mathbb{P}_{\bowtie k}(C_{i\rightarrow j}\varphi)$
   ~~\emph{iff}  $Prob(s \models C_{i\rightarrow j}\varphi) \bowtie \!k$ where: $Prob(s\models C_{i\rightarrow j}\varphi)=\frac{\sum_{s \approx_{i \rightarrow j}s'}|s'\models K_i \varphi \wedge K_j \varphi| }{|s \approx_{i \rightarrow j}s'| };$\\

\noindent $s\models \mathbb{P}_{\bowtie k}(C_{i\rightarrow G}\varphi)$
   ~\emph{iff}  $Prob(s \models C_{i\rightarrow G}\varphi) \bowtie \!k$ where: $Prob(s\models C_{i\rightarrow G}\varphi)=\frac{\sum_{s \approx_{i \rightarrow G}s'}|s'\models K_i \varphi \wedge E_G \varphi| }{|s \approx_{i \rightarrow G}s'| };$\\

\noindent $\bullet$ For a probabilistic operator working over a fulfilment formula, assuming that accessible states are also reachable:\\

\noindent $s\models \mathbb{P}_{\bowtie k}(Fu(C_{i\rightarrow j}\varphi))$
    ~~ \emph{iff}~ $Prob(s \models Fu(C_{i\rightarrow j}\varphi)) \bowtie k;$ where:

\vspace{-0.4cm}
\begin{align*}
%%
& Prob(s\models Fu(C_{i\rightarrow j}\varphi))\ = Prob_s\{\pi \in \Pi(s') ~|~s' \approx_{i \rightarrow j}s ~\textrm{and}~ \pi = s' \ldots s ~\textrm{and}~ s' \models C_{i\rightarrow j}\varphi\};
\end{align*}

\noindent $s\models \mathbb{P}_{\bowtie k}(Fu(C_{i\rightarrow G}\varphi))$
    ~~ \emph{iff}~ $Prob(s \models Fu(C_{i\rightarrow G}\varphi)) \bowtie k;$ where:

\vspace{-0.4cm}
\begin{align*}
%%
& Prob(s\models Fu(C_{i\rightarrow G}\varphi))\ = Prob_s\{\pi \in \Pi(s') ~|~
& s' \approx_{i \rightarrow G}s ~\textrm{and}~ \pi = s' \ldots s ~\textrm{and}~ s' \models C_{i\rightarrow G}\varphi\}
\end{align*}

\end{definition}

Again as in Chapter \ref{cha:PCTLKC}, in the case of the knowledge, the uncertainty is computed in such a way that it reflects the indistinguishability property of the epistemic accessibility relations. Hence, the uncertainty is computed based on the probability of epistemic accessibility relations which is calculated based on the number of accessible states satisfying the content of the knowledge over the number of equivalent states, as all the states are equally accessible. Likewise, probabilistic commitment is computed based on the number of accessible states that satisfy the content over the whole number of accessible states, which demonstrates the uncertainty of the agent over the accessible states, so that over the commitment. Probabilistic fulfillment, however, is computed using the probabilistic transitions of the path linking the commitment state to the fulfillment state.

The following proposition is straightforward from the semantics:

\begin{proposition} \label{proposition}~\\
If $(\mathfrak{M_3},s)\models \mathbb{P}_{\leq0} (Fu(C_{i \rightarrow
j}\varphi))$ and $(\mathfrak{M_3},s)\models Fu(C_{i \rightarrow
j}\varphi)$, then $s$ is not reachable from the commitment state.
\end{proposition}

\begin{theorem} [Epistemic Equivalences]\label{theorm:Epistemic-equivalence} \ \ \ \

>>>>

\chapter{LEarning Chapter}\label{cha:PCTLKC+}

In this chapter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
